A Beginners Guide to Perfomance Factors of Machine Learning Learning Applications with Focus on Neural Nets

### 2017-12-19 paper.tex
changes 19th December

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

\author{\IEEEauthorblockN{Johannes Wünsche}
\IEEEauthorblockA{Otto-von-Guericke University Magdeburg\\
Email: johannes.wuensche@st.ovgu.de}
%\and
%\IEEEauthorblockN{Gabriel Campero Durand}
%\IEEEauthorblockA{Otto-von-Guericke University Magdeburg}
}

@@

\IEEEpeerreviewmaketitle
\section{Introduction}
% no \IEEEPARstart
BE A BIT PITCHIER!\\
Machine Learning seems to everywhere around us. With the success of Deep Learning, Convolutional, Deep Belief, and other Neural Net specifications in the recent years this trend continues to be more and more influencing everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer. But also in modern automation and research machine learning becomes a greater and greater field. We want to have a look at the most performance influencing factors while designing and implementing these networks.

@@

\section{Choice of Calculation Unit}
The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
\subsection{CPU}
As the first most naive option we looked on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
\subsubsection{Single-thread}
The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required \emph{NUMBER OF EXAMPLE CAN BE ADDED HERE}.
\subsubsection{Multi-thread}
A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values. This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm. %maybe a nice little table here with speed improvements of different frameworks over the amount of coures used during the calculation
Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage[link performance paper here(believe it was benchmarking state of the art neural network frameworks)]
\subsection{GPU}
A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs. For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the naive approachs performance. Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.

@@

\section{Communication strategy}
Here I want clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.

@@

\section{Conclusion}
yeah ... pretty much the conclusion shortly repeating the basis that is important for it and point the way they were created, should be larger part to emphasize the importance of some factors over others and the reasoning behind this classification

% use section* for acknowledgment
\section*{Acknowledgment}

The authors would like to thank...

@@

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://mirror.ctan.org/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}
\bibitem{IEEEhowto:kopka}pap,
S. Shi, Q. Wang, P. Xu and X. Chu, \emph{Benchmarking State-of-the-Art Deep Learning Software}, HK Baptist University, Hong Kong, 17th Feb 2017
\end{thebibliography}

>>>>

% *** CITATION PACKAGES ***

\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

@@

\author{\IEEEauthorblockN{Johannes Wünsche}
\IEEEauthorblockA{Otto-von-Guericke University Magdeburg\\
Email: johannes.wuensche@st.ovgu.de}
}

@@

\IEEEpeerreviewmaketitle
\section{Introduction}
% no \IEEEPARstart
Machine Learning seems to be everywhere around us. With the success of Deep , Convolutional, Deep Belief, and other Neural Net specifications in the recent years because of the increasing capabilities of modern hardware, this trend continues to be more and more influencing on everyday application, may it be from music recommendation to personalized advertisement directly addressed to the consumer. But also in modern automation and research machine learning becomes a greater field managing to find more efficient ways to achieve better results in many expertise. We want to have a look at the most performance influencing factors of theses networks while designing and implementing them for specific applications for both small and large scale.

@@

\section{Choice of Processing Unit}
The next important choice is the choice of the actual operating calculation unit. We distinguish them into three main models, CPU, GPU and GPU-Clustering.
\subsection{CPU}
As the first most naive option we looked on the CPU for training of the neural net with the simple single thread and more complex multi-thread calculation.
\subsubsection{Single-thread}
The most naive implementation is the single-thread CPU calculation, because oft this more  brute force approach and sequential execution and calculation this is also the slowest and with a larger number of connected nodes simply unfeasible because of the excessive calculation time required \emph{NUMBER OF EXAMPLE CAN BE ADDED HERE}.
\subsubsection{Multi-thread}
A more complex approach consist of utilizing the multi-core characterisitics of modern CPUs, by splitting of training calculation with only requirements to already calculated values. This is for example preferrable when training with the help of a \emph{Backpropagation} algorithm.
\begin{figure}
\centering
\begin{tabular}{c c c}
FW & 1 Thread & 4 Threads\\
\end{tabular}
\caption{Simulation results for the network.}
\label{fig_sim}
\end{figure}

Though this approach is limited by the actual core number available and should not overstep the number of physical cores, like with intel \emph{hyper-threading} technology, which can lengthen the required calculation time because of a more inefficient usage\cite{shi2016benchmarking}.
\subsection{GPU}
A more advanced implementation is to use the shared memory, multi-core environment of modern GPUs. For example with NVIDIAs CUDA this can be done efficiently and result in a speed-up of calculation of up to 3 times the CPU-based approachs performance for deep learning algorithms over optimized floting-point baseline\cite{shi2016benchmarking}. Allmost all recent neural net learning frameworks support this calculation unit because of its excellent performance rating for neural networks.

@@

\section{Communication strategy}
Here I want to 
clarify a bit on strategies to create an efficient CPU-GPUs interchange with as little overhead as possible.
TODO: Spanish Paper on SuperComputer Learning

@@

\section{Conclusion}
yeah ... pretty much the conclusion shortly repeating the basis that is important for it and point the way they were created, should be larger part to emphasize the importance of some factors over others and the reasoning behind this classification
\cite{shi2016benchmarking}

TODO: Takeaway, different ideas, innnovative

% use section* for acknowledgment
\section*{Acknowledgment}
The author would like to thank Gabriel Campero Durand for advise and help to write the paper

@@

\bibliography{paper}
\bibliographystyle{IEEEtran}
